---
title: "Stat 3301: Homework 1"
date: "Due Wednesday, August 31 by 11:59pm"
author: "Mason Rocco (rocco.58)"
output: html_document
---

\

Setup:
```{r message=FALSE, warning=FALSE}
library(alr4)
library(tidyverse)
```

\



#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions below in the spaces marked "Solution:".
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. Use the option `echo = TRUE` to make sure the R code is displayed.
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
    - formatting all plots to be appropriately sized, with appropriate axis labels.
    - only including R code that is necessary to answer the questions below.
    - only including R output that is necessary to answer the questions below (avoiding lengthy output).
    - providing short written answers explaining your work, and writing in complete sentences.

\

#### Question 1
Import the data set `ames_real_estate.csv` (which is available on Carmen) and use it to do the
following:

a) Find the smallest and largest observed sale prices separately for each neighborhood.
b) Make a scatterplot with the square footage of the house on the *x*-axis and the sale price of the house on the *y*-axis.
c) Write a sentence describing the relationship between `SalePrice` and `SqFt`.

\

##### Solution to Question 1

```{r echo=TRUE}
library(readr)
ames <- read_csv("C:/Users/mrocc/Desktop/Stat 3301/data/ames_real_estate.csv")

ames %>% group_by(ames$Neighborhood) %>% summarize(max(SalePrice))
ames %>% group_by(ames$Neighborhood) %>% summarize(min(SalePrice))

ames %>% ggplot(aes(x=ames$SqFt, y=ames$SalePrice))+geom_point()+
  labs(x="Square Feet", y="Sale Price") + 
  ggtitle("Square Feet vs. Sale Price in Ames, Iowa")
```

It seems that as the square footage of the house increases, the sale price tends
to increase as well.

By using the group by method, I was able to separate the dataset by neighborhood.
Then, I summarized the dataset separately to find the maximum and minimum sale
price in each of the neighborhoods.
\

#### Question 2
Use the data set `Heights` from the library`alr4` to calculate the average height for daughters whose mothers are **at least** 60 inches tall. (The purpose of this question is to make sure you have the package installed correctly.)

\

##### Solution to Question 2
Your answers go here.
```{r eval=FALSE}
heights <- alr4::Heights
heights %>% filter(heights$mheight >= 60) %>%
  summarise(mean=mean(heights$dheight))
```

Using the filter method, I changed the heights dataset to include only the
values with mothers who have heights that are at least 60 inches. I then 
took the average of the heights of the daughters within the heights dataset.
