---
title: "Stat 3301: Homework 2"
date: "Due by September 7, 2022"
author: "Mason Rocco (rocco.58)"
output: html_document
---
\

Setup:
```{r message=FALSE}
library(alr4)
library(tidyverse)
```

\

#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions below in the spaces marked "Solution:".
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. Use the option `echo = TRUE` to make sure the R code is displayed.
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
    - formatting all plots to be appropriately sized, with appropriate axis labels.
    - only including R code that is necessary to answer the questions below.
    - only including R output that is necessary to answer the questions below (avoiding lengthy output).
    - providing short written answers explaining your work, and writing in complete sentences.
- Data files mentioned below are from the `alr4` package unless specified otherwise.

\

#### Question 1
<!-- Basic plotting, simple linear regression relationship, transformations -->
Complete **Problem 1.1** from Weisberg: (Data file: `UN11`) "The data in the file `UN11`..."

**Note**: for part 1.1.3, the `R` function `log` corresponds to the natural logarithm
(base *e*).  The `R` function `log10` corresponds to the base 10 logarithm.  Other bases
can be obtained via `log(x, base = ...)`. Note that in this class we will use the notation
$\log_{10}$ to refer to the base 10 logarithm and $\log$ to refer to the base *e* logarithm
(this is typical in the field of statistics).


\

##### Solution to Question 1

**Part 1.1.1:**

Fertility is the response and PPGDP is the predictor.

**Part 1.1.2:**
```{r, echo=TRUE}
UN <- alr4::UN11
UN %>% ggplot(aes(x=UN$ppgdp, y= UN$fertility)) + geom_point() + 
  labs(x="PPGDP", y="Fertility")
```

I would say that a straight line mean function does not appear to be plausible 
to summarize this graph, since there does not appear to be a linear relationship
between PPGDP and fertility in this graph.

**Part 1.1.3:**
```{r, echo=TRUE}
UN <- alr4::UN11
UN %>% ggplot(aes(x=log(UN$ppgdp), y= log(UN$fertility))) + geom_point() + 
  labs(x="Log of PPGDP", y="Log of Fertility")
```

I would argue that the simple linear regression model is plausible for a summary
of this graph, as it seems there is roughly a negative linear correlation. As
the log of PPGDP increases, the log of fertility decreases.

------


#### Question 2
<!-- Basic plotting, examining relationships between variables, tweaking plots -->
Problem 1.3 from Weisberg: (Data file: `Mitchell`) The data shown in Figure 1.12..."

**Hint**: Part 1.3.2 asks you to make a plot where the length of the horizontal axis is
at least 4 times the length of the vertical axis. This can be done by specifying the
R chunk option `fig.asp = ?` for an appropriate value of the question mark, like this:

<pre><code>```{r fig.asp = ?}
# code goes here
```</code> </pre>

You might try Googling "fig.asp RMarkdown" or something along those lines if you can't
find help on the `fig.asp` option in R. If you get an error that says 
`figure margins too large` when you try to make the plot, just try adjusting the value of `?` 
until you get something close a 4 to 1 ratio.


\

##### Solution to Question 2

**Part 1.3.1:**
It seems as if the soil temperature does not have much dependence on the month
number, as the temperatures are relatively similar for all month numbers.


**Part 1.3.2:**
```{r fig.asp=.33}
mitchell <- alr4::Mitchell
mitchell %>% ggplot(aes(x=Month, y=Temp)) + geom_point() + 
  labs(x="Months After January 1976", y="Average Soil Temperature")
```

After changing the aspect ratio of the graph, it seems as if the temperature of
the soil slightly increases and decreases as the month numbers change, which 
logically makes sense considering how temperatures change with seasons.

------


#### Question 3
<!--Scatterplot matrix, relationships between variables -->
Complete **Problem 1.6** from Weisberg: (Data file `Rateprof`) "In the website and online forum
`RateMyProfessors.com`..."

**Note**: In addition to summarizing the relationships, **also** reproduce the
scatterplot matrix.

Recall that you can create a subset of the whole data set that corresponds to the 
variables you want to plot using:
```{r eval = FALSE}
subset(Rateprof, select = c(quality, clarity, helpfulness, easiness, raterInterest))
```
or
```{r eval = FALSE}
Rateprof %>% select(quality, clarity, helpfulness, easiness, raterInterest)
```


\


##### Solution to Question 3
```{r echo=TRUE}
rate <- alr4::Rateprof
pairs(subset(rate, select= c(quality, clarity, helpfulness, easiness,
                             raterInterest)))
```

When looking at the scatterplot matrix, we can see that there are clearly 
positive linear relationships between quality and clarity, quality and
helpfulness, and clarity and helpfulness. In terms of the other variables, the
relationships are not as clear, but there are loose linear relationships in
most scenarios.
