---
title: "Stat 3301: Homework 4"
date: "Mason Rocco (rocco.58)"
author: "Due by September 23, 2022"
output: html_document
---


\


Setup:
```{r message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(alr4)
library(tidyverse)
```

\

#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions below in the spaces marked "Solution:".
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. I have set the global option `echo = TRUE` to make sure the R code is displayed.
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
    - formatting all plots to be appropriately sized, with appropriate axis labels.
    - only including R code that is necessary to answer the questions below.
    - only including R output that is necessary to answer the questions below (avoiding lengthy output).
    - providing short written answers explaining your work, and writing in complete sentences.
- Data files mentioned below are from the `alr4` package unless specified otherwise.

\

#### Concepts \& Application

In this assignment, you will gain experience with the following methods of statistical inference:

* confidence intervals for mean functions
* hypothesis tests about mean functions
* predictions and prediction intervals
* confidence intervals for functions of parameters


------


#### Questions
<!--
Least squares fit, confidence interval for slope, confidence interval for mean,
confidence bands, raw residuals, interpreting models (comparison)
-->
This is a continuation of the question on the last homework that relates to the **NBA player** 
data we looked at in class (it is available on Carmen).  Here we will focus on players whose 
position is **forward** (labeled `F` in the data set).

Use the statistical model you fit on the last homework assignment to answer the following questions.

1. Calculate and interpret a 95\% confidence interval for the average **weight** of forwards who 
are 81.5\" tall.

2. Remake the plot of weight versus height for forwards with the estimated regression line from the
last homework assignment.  Add lines to the plot representing point-wise 90\% confidence intervals 
for the mean function. Also provide a formula explaining how these intervals are computed.

3. A reporter makes the claim that the average weight for NBA forwards who are 6' 7" tall
is 210 pounds. Use an appropriate hypothesis test to assess whether the there is sufficient
evidence to refute this claim.

4. An NBA team hopes to add a new 6' 7" tall forward to their roster. Predict the weight
(reporting both point and interval predictions) for the weight of this new player.

5. In this part we are interested in the quantity $E( wt \mid ht = x+2) - E(wt \mid ht = x)$
for some baseline height $x$.  We will think of this quantity as a parameter and will call the
parameter $\delta$:
\[
  \delta \equiv E( wt \mid ht = x+2) - E(wt \mid ht = x).
\]
In words, this quantity is the difference in average weight for players that are two inches
apart in height.

     a. Show that, in terms of the model parameters, $\delta = 2\beta_1$.
     b. Calculate an estimate $\widehat{\delta} = 2\widehat{\beta}_1$ of $\delta$ using the OLS estimate $\widehat{\beta}_1$.
     c. Find a formula (not a numeric value) for the variance of the estimator:
     $\mbox{Var}(\widehat{\delta} \mid X) = \mbox{Var}(2\widehat{\beta}_1 \mid X) = \ldots$
     d. Find a formula (not a numeric value) for the standard error of the estimator, $se(\widehat{\delta} \mid X)$, by taking the square root of your answer from part (c) and substituting the estimate $\widehat{\sigma}$ for the unknown parameter $\sigma$.
     e. Use the data to calculate the numeric value of $se(\widehat{\delta} \mid X)$.
     f. Use all of the above information to calculate a 95\% confidence interval for $\delta$ and write a sentence interpreting this interval in terms of NBA forwards.
     

\

##### Solutions

**Part 1: **
```{r}
library(readr)
nbahtwt <- read_csv("C:/Users/mrocc/Desktop/Stat 3301/data/nbahtwt.csv")
forwards <- subset(nbahtwt, pos == "F")
forwards.lm <- lm(wt ~ ht, data = forwards)
predict(forwards.lm, newdata = data.frame(ht = 81.5), 
        se.fit = TRUE, interval = "confidence")
```

As can be seen above, the 95% confidence interval for the average weight of forwards who are 81.5" tall is (231.4157, 236.5036).

**Part 2: **
```{r}
hts.grid = seq(from = min(forwards$ht), to = max(forwards$ht), 
               length.out = 100)

forwards.cis <- predict(forwards.lm, newdata = data.frame(ht = hts.grid), 
                        interval = "confidence", level = .90)

plot(forwards$ht, forwards$wt, xlab="Height", ylab="Weight")
lines(hts.grid, forwards.cis[, "fit"], lwd = 1.5)
lines(hts.grid, forwards.cis[, "lwr"], col = "red")
lines(hts.grid, forwards.cis[, "upr"], col = "red")
```

The 90% joint confidence region for the mean function is calculated using the 
function $\hat{\mu}_{Y|X=x} \pm[2F(\alpha;2, n-2)]^{1/2}sefit(\hat{y}|x)$

**Part 3: **
\
$H_0: E(Y|X = 79) = 210$
\
$H_1: E(Y|X = 79) \neq 210$

```{r}
n <- length(forwards$ht)
estimate <- predict(forwards.lm, newdata = data.frame(ht=(79)), se.fit = TRUE)
t <- (estimate$fit-210)/estimate$se.fit
2*pt(abs(t), df= n-2, lower.tail = FALSE)
```

Since the p-value for this proposed hypothesis test is lower than any reasonable
$\alpha$ value, we reject $H_0$. We have significant evidence that the average
weight for NBA forwards who are 6'7" tall is not 210 pounds.

**Part 4: **
```{r}
predict(forwards.lm, newdata = data.frame(ht=(79)),
        interval = "confidence")
predict(forwards.lm, newdata = data.frame(ht=(79)),
        interval = "prediction")
```

If an NBA team was to add a 6'7" forward to their roster, the point estimate for
their weight would be 224.52 pounds, while a 95% confidence interval would be
(221.37, 227.66), and a 95% prediction interval would be (192.99, 256.05).

**Part 5a: **
\
Given the formula $E(Y \mid X = x+c) - E(Y \mid X = x) = c\beta_1$, we can infer
that $\delta = 2\beta_1$.

**Part 5b: **
\
$\hat{\delta} = 2\hat{\beta_1} = 2(\frac{SXY}{SXX})$
```{r}
SXY <- with(forwards, sum((ht- mean(ht)) * (wt- mean(wt))))
SXX <- with(forwards, sum((ht- mean(ht))^2))
b1hat <- SXY/SXX
deltahat <- 2 * b1hat
print(deltahat)
```



**Part 5c: **
\
$Var(\beta_1|X) = \frac{\sigma^2}{SXX}$, $Var(2\hat{\beta_1}|X) = 4\frac{\sigma^2}{SXX}$

**Part 5d: **
\
$se(\beta_1|X) = \frac{\sigma^2}{\sqrt{SXX}}$, $se(\hat\delta|X) =  4\frac{\sigma^2}{\sqrt{SXX}}$

**Part 5e: **
```{r}
se <- (4*var(forwards$ht))/sqrt(SXX)
print(se)
```

**Part 5f: **
```{r}
t <- qt(p= .95, df= n - 2)
lwr <- deltahat - (t * se)
upr <- deltahat + (t * se)
CI <- c(lwr, upr)
print(CI)
```

With this interval, we can say that for NBA forwards, a 2 inch difference in 
height is estimated to be associated with a difference in average weight between
6.87 and 8.24 pounds, with a 95% level of confidence.

------


