---
title: "Stat 3301: Homework 5"
date: "Mason Rocco (rocco.58)"
author: "Due by October 3, 2022"
output: html_document
---


\


Setup:
```{r message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(alr4)
library(tidyverse)
library(broom)
```

\

#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions below in the spaces marked "Solution:".
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. I have set the global option `echo = TRUE` to make sure the R code is displayed.
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
    - formatting all plots to be appropriately sized, with appropriate axis labels.
    - only including R code that is necessary to answer the questions below.
    - only including R output that is necessary to answer the questions below (avoiding lengthy output).
    - providing short written answers explaining your work, and writing in complete sentences.
- Data files mentioned below are from the `alr4` package unless specified otherwise.

\

#### Concepts \& Application

In this assignment, you will

* calculate and interpret an $R^2$ statistic.
* use appropriate residuals to check the normal SLR model assumptions.
* fit regression models with the predictor in log scale.
* interpret $\hat{\beta}_1$ when the predictor is in log scale.
* construct confidence intervals when the predictor is in log scale.
* make a plot of a fitted model when the predictor is in log scale.

------

#### Question 1
<!--
R2, residual diagnostic plots
-->
This is continuation of the question on the last homework that relates to the **NBA player** 
data we looked at in class (it is available on Carmen).  Here we will focus on players whose 
position is **forward** (labeled `F` in the data set).

Use the statistical model you fit on the last homework assignment to answer the following questions.

1. Calculate the total sum of squares, the regression sum of squares and the residual sum
of squares for the fitted regression model.  Use these to calculate the coefficient of
determination for the model. What do these values tell you about the
strength of the linear relationship between weight and height for NBA forwards?

2. Use appropriate residual plots to assess whether the fitted mean function is appropriate.
Do the plots indicate any lack of fit?

3. Use appropriate residual plots to assess the assumption that the variance of player
weight is the same at all heights.

4. Use appropriate residual plots to assess the assumption that the error terms in the model
are normally distributed. 


\


##### Solution to Question 1
\

### 1
```{r}
library(readr)
nbahtwt <- read_csv("C:/Users/mrocc/Desktop/Stat 3301/data/nbahtwt.csv")
forwards <- subset(nbahtwt, pos == "F")
forwards.lm <- lm(wt ~ ht, data = forwards)
RSS <- sum(resid(forwards.lm)^2)
SSreg <- sum((fitted(forwards.lm) - mean(forwards$wt))^2)
SYY <- RSS + SSreg
R2 <- SSreg/SYY
print(RSS)
print(SSreg)
print(SYY)
print(R2)
```

When looking at the value of the coefficient of determination, we can see that
there is not a very strong linear relationship between the weight and height for
NBA forwards, since the value is 0.1119, which is relatively low.

### 2
```{r}
(eplot = augment(forwards.lm) %>% ggplot(aes(x=.fitted, y=.resid)) +
   geom_point(size = 2) + geom_hline(yintercept=0) + theme_bw(22) + 
   xlab("Fitted Values") + ylab("Residuals") + geom_smooth(se=FALSE))
```

This plot seems to be relatively linearly distributed, so there is no clear 
indication of the mean function not being appropriate.

### 3
```{r}
plot(fitted(forwards.lm), rstandard(forwards.lm),
     xlab = "Fitted Values", 
     ylab = "Standardized Residuals"); abline(h=0)
```

This plot shows that the variance of player weight is mostly the same at all
heights.

### 4
```{r}
qqnorm(resid(forwards.lm)); qqline(resid(forwards.lm))
hist(resid(forwards.lm), xlab = "Residuals")
```

When looking at these residual plots, we can see that the error terms in the
model are normally distributed, as there is no strong evidence against normality.

------


#### Question 2
<!--
...
-->
The data set `ufcwc` in the `alr4` package contains data on a sample of western cedar trees
that was taken in 1991 in the Upper Flat Creek stand of the University of Idaho
Experimental Forest. Each case in the data set represents a tree. In this question we will 
look at the relationship between the variables:

| Variable      | Description                                                    |
| ------------- | -------------------------------------------------------------- |
| `Height`      | height of the tree in decimeters                               |
| `Dbh`         | diameter of the tree in mm measured at 137 cm above the ground |

We will be interested in using tree diameter (`Dbh`) to predict tree height (`Height`).

(a) For the linear regression model $Height_i = \beta_0 + \beta_1 \log Dbh_i + e_i$, report
the numeric values of estimated parameters $\hat{\beta}_0$, $\hat{\beta}_1$ and $\hat{\sigma}^2$.

(b) Make a scatterplot with `log(Dbh)` on the x-axis and `Height` on the y-axis, and add
the estimated linear regression line to the plot.

(c) Make a scatterplot with `Dbh` (on its original scale) on the x-axis and `Height` on 
the y-axis. Use your fitted model from part (a) to add a (curved) line to the plot that represents 
the estimated average tree height as a function of tree diameter, $\hat{E}(\mathtt{Height} \mid 
\mathtt{Dbh})$. Add dashed (curved) lines to the plot that represent point-wise, 95\% confidence
intervals for the mean.

(d) Using the estimated slope coefficient, $\hat{\beta}_1$, write a sentence that quantifies
what our model says about how average tree height depends on tree diameter (on its original scale).



\

##### Solution to Question 2
\

### a
```{r}
library(alr4)
trees <- alr4::ufcwc
treemod <- lm(Height ~ log10(Dbh), data = trees)
treesum <- summary(treemod)
treemod
treesum$sigma^2
```

For this linear regression model, $\hat\beta_0 = -463.3$,
$\hat\beta_1 = 275.2$, and $\hat\sigma^2 = 1111.185$.

### b
```{r}
trees%>%ggplot(aes(x = log(Dbh), y = Height)) + geom_point() + 
    geom_smooth(method = "lm", se = F)
```

### c
```{r}
dbh.grid <- data.frame(Dbh=seq(from=min(trees$Dbh), to=max(trees$Dbh), 
                                          length.out=139))
ypred <- predict(treemod, newdata = dbh.grid, interval = "confidence", level = .95)
plot(trees$Dbh, trees$Height, xlab = "Tree Diameter", ylab = "Tree Height",
     main = "Estimated Mean Function with 95% Confidence Interval")
lines(dbh.grid[,1], ypred[,1], col="blue")
lines(dbh.grid[,1], ypred[,2], col="blue", lty=2)
lines(dbh.grid[,1], ypred[,3], col="blue", lty=2)
```


### d
```{r}
betahat_1 <- 275.2
betahat_1 * log(1.1)
```

Since we know that $\hat\beta_1 = 275.2$, we can say that a 10% increase in height
is associated with an additional 26.23mm of diameter for a western cedar tree in
the Upper Flat Creek stand of the University of Idaho Experimental Forest.

------
