---
title: "Stat 3301: Homework 6"
date: "Mason Rocco (rocco.58)"
author: "Due by October 19, 2022"
output: html_document
---


\


Setup:
```{r message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(alr4)
library(tidyverse)
library(readr)
library(broom)
library(GGally)
```

\

#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions below in the spaces marked "Solution:".
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. I have set the global option `echo = TRUE` to make sure the R code is displayed.
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
    - formatting all plots to be appropriately sized, with appropriate axis labels.
    - only including R code that is necessary to answer the questions below.
    - only including R output that is necessary to answer the questions below (avoiding lengthy output).
    - providing short written answers explaining your work, and writing in complete sentences.
- Data files mentioned below are from the `alr4` package unless specified otherwise.

\

#### Concepts \& Application

In this assignment, you will

* plot pairwise relationships between three variables.
* fit a multiple linear regression model in R.
* gain experience with several matrix operations in R.
* calculate a "hat value" (leverage) for a case.
* use plots of standardized residuals to assess model fit.

------

#### Question 1

This question is concerned with the following variables from the NBA data set for **guards** only:

| Variable  | Description                            |
|:--------- |:-------------------------------------- |
| `wt`      | weight of player (guard) in pounds     |
| `ht`      | height of player (guard) in inches     |
| `age`     | age of player (guard) in years         |                  

We will be interested in using `ht` and `age` jointly to predict `wt` for **guards** (ignoring the other positions for now).

(a) Make a scatterplot matrix for the variables `ht`, `age` and `wt`. Use the plots to describe the pairwise relationships between the variables.

(b) Use `R` to fit a multiple linear regression model with mean function $E(\mathtt{wt} \mid
\mathtt{ht}, \mathtt{age}) = \beta_0 + \beta_1 \mathtt{ht} + \beta_2 \mathtt{age}$ and
variance function $\mbox{Var}(\mathtt{wt} \mid \mathtt{ht}, \mathtt{age}) = \sigma^2$.
Use the `summary` function to display `R`'s standard summary of the fitted model.

(c) Report the numeric values of the parameter estimates $\hat{\beta}_j$, $j = 0, 1, 2$, 
and $\hat{\sigma}^2$.

(d) Report the numeric value of the degrees of freedom associated with the fitted model. Say what
formula you use to calculate this value.

(e) Use R to compute and display the 3x3 matrix $X^T X$, where *X* is the $202 \times 3$ matrix used to fit the regression model.

(f) Use R to compute and report the 3x3 matrix $(X^TX)^{-1}$.

(g) Nate Robinson should be the first case in the data set ($i = 1$). Compute and report 
the leverage $h_{11}$ for Nate Robinson under this MLR model in two ways: (i) using the 
`hatvalues` function and (ii) by computing the **H** matrix directly and reporting the 
first diagonal element.

(h) Make three plots of the standardized residuals $r_i$: (i) $r_i$ versus the fitted values
$\hat{y}_i$, (ii) $r_i$ versus `ht` and (iii) $r_i$ versus `age`. Do the plots suggest any
problems with our assumptions about the mean and variance functions?

(Note: you will notice some diagonal "stripes" in the plot of residuals versus fitted values.
These are due to the fact that the weights are rounded to the nearest pound, and are **not** 
a cause for concern.)



\


### Solution to Question 1
#### a
```{r}
nbahtwt <- read_csv("C:/Users/mrocc/Desktop/Stat 3301/data/nbahtwt.csv")
guards <- nbahtwt %>% filter(pos == "G")
guards %>% ggpairs(columns = c("ht", "wt", "age"))
```

We can see there is a relatively strong positive linear relationship between 
weight and height. That is, as height increases for NBA guards, the weight of
NBA guards tends to increase as well. There are not strong linear relationships
between weight and age or height and age of NBA guards. 

#### b
```{r}
guards.lm <- lm(wt ~ ht + age, data=guards)
guards.summary <- summary(guards.lm)
print(guards.summary)
```

#### c
```{r}
guards.summary$coefficients
(guards.summary$sigma)^2
```

We can see that $\hat\beta_0$ is -148.82, $\hat\beta_1$ is 4.43, and
$\hat\beta_2$ is 0.41. Additionally, $\hat\sigma^2$ is 142.97.

#### d
The degrees of freedom for the fitted model is 199. This can be calculated using
the formula n-(p+1) = 202 - (2+1) = 199. 

#### e
```{r}
X <- model.matrix(guards.lm)
XtX <- (t(X)) %*% X
print(XtX)
```

#### f
```{r}
solve(XtX)
```

#### g
##### i
```{r}
hatvalues(guards.lm)[1]
```

##### ii
```{r}
H <- X %*% solve(t(X)%*%X)%*%t(X)
H[1]
```

#### h
##### i
```{r}
plot(fitted(guards.lm), rstandard(guards.lm), xlab = "Fitted Values", 
     ylab = "Standardized Residuals")
```

##### ii
```{r}
plot(guards$ht, rstandard(guards.lm), xlab = "Height", 
     ylab = "Standardized Residuals")
```

##### iii
```{r}
plot(guards$age, rstandard(guards.lm), xlab = "Age", 
     ylab = "Standardized Residuals")
```

When looking at the three plots of the standardized residuals, we can see that
none of the plots suggest any problems with our assumptions about the mean and
variance functions. This is true because the plots all appear to be null plots
with no fanning patterns, suggesting that the mean and variance and normal.

------
