---
title: "Stat 3301: Homework 7"
date: "Mason Rocco (rocco.58)"
author: "Due by October 26, 2022"
output: html_document
---


\


Setup:
```{r message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(alr4)
library(tidyverse)
```

\


#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions below in the spaces marked "Solution:".
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. I have set the global option `echo = TRUE` to make sure the R code is displayed.
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
    - formatting all plots to be appropriately sized, with appropriate axis labels.
    - only including R code that is necessary to answer the questions below.
    - only including R output that is necessary to answer the questions below (avoiding lengthy output).
    - providing short written answers explaining your work, and writing in complete sentences.
- Data files mentioned below are from the `alr4` package unless specified otherwise.

\

#### Concepts \& Application

In this assignment, you will gain experience with

* calculating the covariance matrix for the estimated regression coefficients.
* calculating various standard errors.
* interpreting coefficients in a MLR model.
* performing hypothesis tests about individual coefficients.
* computing confidence intervals for the mean of the response.

------

#### Question 1

This question is concerned with the following variables from the NBA data set for **guards** only:

| Variable      | Description                            |
| ------------- | -------------------------------------- |
| `wt`          | weight of player (guard) in pounds     |
| `ht`          | height of player (guard) in inches     |
| `age`         | age of player (guard) in years         |                  

We will be interested in using `ht` and `age` jointly to predict `wt` for **guards** (ignoring the other positions for now) using the linear regression model
\[
  wt_i = \beta_0 + \beta_1 ht_i + \beta_2 age_i + e_i,
\]
where $e_i \stackrel{\mathrm{iid}}{\sim}\mbox{N}(0, \sigma^2)$, $i = 1, \ldots, n$. A summary of the
fitted regression model is

```{r, echo = FALSE, message = FALSE}
guards = read_csv("C:/Users/mrocc/Desktop/Stat 3301/data/nbahtwt.csv") %>% filter(pos == "G")
summary(lm(wt ~ ht + age, data = guards))
```


(a) Use the **X** matrix used to fit the model along with $\hat{\sigma}$ to compute the estimated covariance
matrix for the estimated coefficient vector, 
$\widehat{\mbox{Var}}(\hat{\boldsymbol \beta} \mid X)$. Show your code and computations. Verify your
result by comparing your answer to the result obtained using the `vcov` function.

(b) Use your result from part (a) to calculate the standard error for the estimated coefficient for the
regressor `ht`. Verify your answer by comparing it with the summary of the fitted regression model 
displayed above.

(c) The estimated coefficient for height is $\hat{\beta}_1 = 4.4250$. Write a sentence interpreting
this numeric value in terms of the relationship between height and weight.

(d) Now focusing on the coefficient for age, the `t value` displayed in the output above is 2.072. 
Write down the hypotheses this test statistic can be used to test using appropriate statistical notation.
Use the output above to make a conclusion (write a sentence describing the result).

(e) Say we are interested in estimating the average weight for 20-year-old guards who are 78.5" tall. Compute the standard error needed to compute a confidence interval for this quantity **without** using the `predict` function. Confirm your answer using the `predict` function.

(f) Use your result from (e) to compute a 90\% confidence interval for the average weight for 20-year-old 
guards who are 78.5" tall. Confirm your answer using the `predict` function, and write a sentence 
interpreting the confidence interval in words.



\


#### Solution to Question 1
##### a
```{r}
library(readr)
nbahtwt <- read_csv("C:/Users/mrocc/Desktop/Stat 3301/data/nbahtwt.csv")
guards <- nbahtwt %>% filter(pos == "G")
guards.mod <- lm(wt ~ ht + age, data = guards)
X <- model.matrix(guards.mod)
sig2hat <- summary(guards.mod)$sigma^2
cov.matrix <- sig2hat * solve(t(X)%*%X)
vcov(guards.mod)
print(cov.matrix)
```

##### b
```{r}
sqrt(diag(vcov(guards.mod)))
summary(guards.mod)$coefficients[,2]
```

##### c
Given that the value of $\hat\beta_1$ is 4.4250, we can say that as the height
of NBA guards increases by 1 inch, the weight of NBA guards tends to increase by
4.425 pounds.

##### d
$H_0:\hat\beta_2=0$
$H_1:\hat\beta_2\neq0$
Since the p-value for the test is .03, less than 0.05 (a reasonable $\alpha$ value)
we can say that we will reject the null hypothesis that $\hat\beta_2=0$. This tells
us that $\beta_2$ is a useful predictor after adjusting for the other regressors
in this model.

##### e
```{r}
Y <- matrix(cbind(rep(1, 202), guards$age, guards$ht), ncol = 3)
YtYinv <- solve(t(Y)%*%Y)

ystar <- matrix(c(1, 20, 78.5), ncol = 1)
tmp <- t(ystar)%*%YtYinv%*%ystar
sighat <- summary(guards.mod)$sigma
sefit <- sighat * sqrt(tmp)
print(sefit)
print(predict(guards.mod, newdata = data.frame(ht=78.5, age=20), 
              interval = "confidence", se.fit = TRUE)$se.fit)
```

##### f
```{r}
(-148.8194 + (4.425*78.5) + (.4146*20)) + c(-1,1) * qt(.9, df=199)*c(sefit)

predict(guards.mod, newdata = data.frame(ht=78.5, age=20), 
              interval = "confidence", level = .9)
```

We can see that the 90% confidence interval for the average weight for 20 year 
old guards who are 78.5" tall is *(204.60, 209.07)*. This tells us that we are
90% confident that a 20 year old guard who is 78.5" tall would weigh between
204.60 and 209.07 pounds.

------
