---
title: "Stat 3301: Homework 8"
date: "FirstName LastName (name.n)"
author: "Due by date and time specified on Carmen"
output: html_document
---


\


Setup:
```{r message = FALSE}
library(alr4)
```

\

#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions below in the spaces marked "Solution:".
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. I have set the global option `echo = TRUE` to make sure the R code is displayed.
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
    - formatting all plots to be appropriately sized, with appropriate axis labels.
    - only including R code that is necessary to answer the questions below.
    - only including R output that is necessary to answer the questions below (avoiding lengthy output).
    - providing short written answers explaining your work, and writing in complete sentences.
- Data files mentioned below are from the `alr4` package unless specified otherwise.

\

#### Concepts \& Application

In this assignment, you will gain experience with

* creating dummy variables.
* interpreting coefficients in a one-factor regression model.
* identifying the mean function for a one-factor model based on R output.
* constructing confidence intervals for group means.
* performing hypothesis tests about differences in group means.


#### Question 1

This question concerns the `UN11` data set in the textbook. We will use the factor variable `group`
to explain differences in the distribution of `log(ppgdp)`, the natural logarithm of the `ppgdp`
variable. See the textbook and `?UN11` for more information about the data set. For this question,
`group` is the predictor variable and `log(ppgdp)` is the response variable.


(a) Consider the linear model that uses `group` to explain `log(ppgdp)`:
$$
  \log(\mathtt{ppgdp}_i) = \beta_0 + \beta_1 U_{oecd, i} + \beta_2 U_{africa, i} + e_i,
$$
where $e_i \stackrel{\mathrm{iid}}{\sim} \mbox{N}(0, \sigma^2)$ and the *U* variables
are dummary variables (0/1 indicators) for the value of the `group` variable. Manually
construct the dummy variables $U_{oecd,i}$ and $U_{africa,i}$ needed to fit the model.
Show your code.

(b) Fit the model from part (b) using the dummy variables you constructed. Report the values
of the estimated parameters $\hat{\beta}_0$, $\hat{\beta}_1$, $\hat{\beta}_2$ and $\hat{\sigma}^2$.

(c) For each of the estimated parameters $\hat{\beta}_0$, $\hat{\beta}_1$ and $\hat{\beta}_2$, write
a sentence interpreting what the coefficient means in terms of the relationship between `log(ppgdp)`
and the various levels of the `group` variable.

(d) Show how to use the coefficients $\hat{\boldsymbol{\beta}}$ you estimated above to calculate an 
estimate of the mean log per person gdp for `oecd` countries, $\hat{E}(\mathtt{log(ppgdp)} \mid 
\mathtt{group} = \mathtt{oecd})$.

(e) Show how to use the coefficients $\hat{\boldsymbol{\beta}}$ you estimated above to calculate an 
estimate of the difference in mean log per person gdp bewteen `oecd` countries and countries in `africa`,
$\hat{E}(\mathtt{log(ppgdp)} \mid \mathtt{group} = \mathtt{oecd}) -\hat{E}(\mathtt{log(ppgdp)} \mid \mathtt{group} = \mathtt{africa})$.

```{r}
summary(lm(log(ppgdp) ~ group, data = UN11))
vcov(lm(log(ppgdp) ~ group, data = UN11))
```


Now we allow R to fit the model using the `group` variable directly. You will use the following output to answer the remaining questions.

(f) Use the output above to construct a 95% confidence interval for the mean log per person GDP
for OECD countries. Do not use the `predict` function; show how to use the output above to
compute the confidence interval.

(g) Use the output above to perform a hypothesis test to assess whether the mean log per person GDP in countries 
in Africa is the same as the mean log per person GDP for countries in the "other" category. Clearly write down 
your hypotheses in terms of model **parameters** ($\beta$s), show how you calculate the test statistic, provide 
a *p*-value and make a conclusion.

\


## Solution to Question 1
### a
```{r}
UN11 <- alr4::UN11
groupoecd <- as.numeric(UN11$group == "oecd")
groupafrica <- as.numeric(UN11$group == "africa")
```

### b
```{r}
UN.mod <- lm(log(UN11$ppgdp) ~ groupoecd + groupafrica, data = UN11)
summary(UN.mod)$coefficients
summary(UN.mod)$sigma^2
```

$\hat\beta_0 = 8.60$
\
$\hat\beta_1 = 1.76$
\
$\hat\beta_2 = -1.55$

### c
$\hat\beta_0 = 8.60$, meaning that countries that are categorized as other have
an expected mean of log of ppgdp that is 8.60.
\
$\hat\beta_1 = 1.76$, meaning that countries that are members of the OECD have an
expected mean of log of ppgdp that is 1.76 higher than the baseline, countries
that are categorized as other.
\
$\hat\beta_2 = -1.55$, meaning that countries that are in Africa have an expected
mean of log of ppgdp that is 1.55 lower than the baseline, countries that are
categorized as other.

### d

$\hat{E}(log(ppgdp)|group=oecd) = \hat\beta_0 + \hat\beta_1$
$\hat{E}(log(ppgdp)|group=oecd) = 10.36$

### e
$\hat{E}(log(ppgdp)|group=oecd) - \hat{E}(log(ppgdp)|group=africa)$
$=(\hat\beta_0 + \hat\beta_1) - (\hat\beta_0 + \hat\beta_2)$
$=(\hat\beta_1) - (\hat\beta_2) = 3.31$

### f
```{r}
group.mod <- (lm(log(ppgdp) ~ group, data = UN11))
se <- summary(group.mod)$coefficients[1,2]
t <- qt(.95, 196)
Bhat0 <- summary(group.mod)$coefficients[1,1]
lwr <- Bhat0 - t*se
upr <- Bhat0 + t*se
print(c(lwr,upr))
```

### g
$H_0: \beta_1 - \beta_2 = 0$
\
$H_1: \beta_1 - \beta_2 \neq 0$
\
$t = \frac{(\beta_1 - \beta_2)-0}{se(\beta_1 - \beta_2)}$
\
```{r}
tvalue <- (3.3059 - 1.7578) / (0.05463725 + 0.06820830 -2*(0.04303619))
2 * pt(tvalue, df=196, lower.tail = F)
```

Given that the p-value for this hypothesis test is very low, we reject the null
hypothesis that $\beta_1 - \beta_2 = 0$. This means that we do not have 
statistically significant evidence to say that the mean log(ppgdp) of African 
countries is equal to the mean log(ppgdp) of countries in the "other" category.


------




