---
title: "Stat 3301: Homework 9"
date: "FirstName LastName (name.n)"
author: "Due by date and time specified on Carmen"
output: html_document
---


\


Setup:
```{r message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(alr4)
library(tidyverse)
library(broom)
```

\

#### Instructions
- Replace "FirstName LastName (name.n)" above with your information.
- Provide your solutions below in the spaces marked "Solution:".
- Include any R code that you use to answer the questions; if a numeric answer is required, show how you calculated it in R. I have set the global option `echo = TRUE` to make sure the R code is displayed.
- Knit this document to HTML and upload both the HTML file and your completed Rmd file to Carmen
- Make sure your solutions are clean and easy-to-read by
    - formatting all plots to be appropriately sized, with appropriate axis labels.
    - only including R code that is necessary to answer the questions below.
    - only including R output that is necessary to answer the questions below (avoiding lengthy output).
    - providing short written answers explaining your work, and writing in complete sentences.
- Data files mentioned below are from the `alr4` package unless specified otherwise.

\

#### Concepts \& Application

In this assignment, you will gain experience with

* construct confidence intervals for the unknown mean of a response variable in a model with one categorical predictor and one continuous predictor.
* fitting models with an interaction between a continuous and a categorical predictor.
* plotting and comparing fitted models.

------

## Question 1
This question concerns the `UN11` data set in the textbook that we used in the previous homework. Here we consider a "main effects" model that uses both `log(ppgdp)` and `group` to predict female life expectancy (`lifeExpF`). The "main effects" model allows for different intercepts for the regression of `lifeExpF` on `log(ppgdp)` for different levels of the `group` variable but requires that the slope parameters are the same across levels of `group`. Use the R output below to answer the following questions.
```{r}
summary(lm(lifeExpF ~ group + log(ppgdp), data = UN11))
```

(a) Use appropriate dummy variables to write down the mean function that R used when fitting the model above:
\[
  E(lifeExpF \mid group, log(ppgdp)) = \hspace{4in}
\]


(b) Construct a 95% confidence interval for the coefficient labeled `groupafrica` and explain its meaning.

(c) Explain the meaning of the estimated coefficient for `log(ppgdp)`.

(d) Use the `predict` function to compute a 95% confidence interval for the average female life expectancy
for countries in the OECD group with a **log** per person gross domestic product of 10.5 log US dollars per person.

(e) Explain what hypotheses are being tested in the `groupother` row of the output above. Write a sentence 
interpreting the result of the test.



\


## Solution to Question 1
### a
  \[E(lifeExpF \mid group, log(ppgdp)) = \beta_0+\beta_1(U_{other}i)
  +\beta_2(U_{africa}i)+\beta_3(log(ppgdp)_i)+e_i, 
  e_i \stackrel{\mathrm{iid}}{\sim} \mbox{N}(0, \sigma^2)
  \]
  
### b
```{r}
UN <- alr4::UN11
UN.lm <- lm(lifeExpF ~ group + log(ppgdp), data = UN)
Bhat0 <- summary(UN.lm)$coefficient[3,1]
se <- summary(UN.lm)$coefficient[3,2]
t <- qt(.95/2, 195)
lwr <- Bhat0 - t*se
upr <- Bhat0 + t*se
print(c(lwr,upr))
```

Given that the 95% confidence interval for $\beta_2$ is **(-12.07, -12.27)**, we
can say that we are 95% confident that the value of $\beta_2$ is captured by the
interval.

### c
Given that the estimated coefficient for log(ppgdp) is 3.177, we can say that
for females in the same group, a 10% increase in ppgdp is associated with an
estimated 30.28% increase in years of life expectancy.

### d
```{r}
predict(UN.lm, newdata = data.frame(group = "oecd", ppgdp = exp(10.5)), 
        level = .95, interval = "pred")
```

### e
In the above output, the hypotheses being tested are as follows:
\
$H_0: \beta_1 = 0$
\
$H_1: \beta_1 \neq 0$
\
This means that we are trying to see if the estimate for the coefficient of
groupother is 0 or not. As the p-value is 0.193, which is higher than most
reasonable $\alpha$ values, we will fail to reject the null hypothesis. This means
that we do not have enough evidence to say that the coefficient for groupother is
not 0.
\
------

## Question 2
**NHL Data Revisited**

Earlier this semester in class we considered the model
$$
  \log ttsm_i = \beta_0 + \beta_1 \log estFans_i + e_i
$$
that used the estimated number of NHL fans in a city to predict total ticket spending. In this version of the model, $\log$ refers to base $e$ logarithm, and `ttsm` is the total ticket spending in millions of dollars. The fitted model looked like:

```{r echo=FALSE, fig.height=3, fig.width=4, fig.align = "center", message = FALSE}
nhl = read.csv("C:/Users/mrocc/Desktop/Stat 3301/data/nhlfans3301.csv", header=T) %>% 
  mutate(ttsm = tts/1e6)
nhl %>% ggplot(aes(x=log(estFans), y=log(ttsm), color=country)) + geom_point() + geom_smooth(method="lm", se=F, aes(color=NA), size=0.75)
```

Our analysis earlier in the semester ignored the variable `country`. We will now consider whether `country` is a useful predictor in the model after adjusting for the effect of estimated fans. Another way to think about this is whether the relationship between `ttsm` and `estFans` is different for teams in the USA vs. Canada (and if so, how is that relationship different).

(a) Fit a model that includes both main and interaction effects for `log(estFans)` and `country` (i.e., a model that allows for different intercepts and slopes for `log(estFans)` across different levels of `country`). Write down the estimated regression equations separately for teams in the USA and Canada. Make a plot of the data (color coded by country) that includes the two estimated regression lines (color coded by country).

(b) Do the same as you did in part (a), but now for the model that includes **only** main effects terms for both predictors (i.e., a model with the same slope but different intercepts for the different levels of `country`).

(c) You now have three potential models: one that assumes the relationship between `log(ttsm)` and `log(estFans)` is exactly the same no matter the country (`log(ttsm) ~ log(estFans)`), one that assumes parallel regression lines (`log(ttsm) ~ country + log(estFans)`), and one that allows for different regression relationships (`log(ttsm) ~ country*log(estFans)`). Based on the plots and the summaries of the fitted models (*t*-tests, etc.), which model do you think best describes the relationship between variables?



\


## Solution to Question 2
Your answers go here.

### a
```{r}
nhl <- read_csv("C:/Users/mrocc/Desktop/Stat 3301/data/nhlfans3301.csv")
modInt <- lm(log(tts/1e6) ~ country * log(estFans), data = nhl)
plotInt = augment(modInt) %>% 
  ggplot(aes(x = log(nhl$estFans), y = log(nhl$tts/1e6), color = nhl$country)) + geom_point() +
  geom_line(aes(y = .fitted), size = 1) + theme_bw(22) +
  xlab("log(estFans)") + ylab("log(ttsm)") + labs(color = "Country")
plotInt
```


### b
```{r}
modMain <- lm(log(tts/1e6) ~ country + log(estFans), data = nhl)
plotMain = augment(modMain) %>% 
  ggplot(aes(x = log(nhl$estFans), y = log(nhl$tts/1e6), color = nhl$country)) + geom_point() +
  geom_line(aes(y = .fitted), size = 1) + theme_bw(22) +
  xlab("log(estFans)") + ylab("log(ttsm)") + labs(color = "Country")
plotMain
```


### c
```{r}
summary(modInt)
summary(modMain)
```

When looking at the three potential models for the relationship between log(ttsm)
and log(estFans) while considering the country of the team, I would say that the
model that best describes the relationships between the variables is the model
that allows for different regression relationships. The reason for this is that
this model seems to more accurately shows the slopes and intercepts, and the
t-tests and p-values reflect that.
